---
title: "Analyses for latent-network/confirmatory factor analysis/ Exploratory graph analysis paper"
author: "Or Duek"
output:
html_document:
df_print: paged
---

## 1. Load libraries
```{r}
# Data handeling
# if(!require("tidyverse")) install.packages("tidyverse")
# if(!require("corrplot")) install.packages("corrplot")  ## correlation matrix plots
# if(!require("OpenMX")) install.packages("OpenMx") 
# 
# # Network packages
 if(!require("qgraph")) install.packages("qgraph")
 if(!require("psychonetrics")) install.packages("psychonetrics")
 if(!require("bootnet")) install.packages("bootnet")
 if(!require("mgm")) install.packages("mgm")
 if(!require("networktools")) install.packages("networktools")
 if(!require("EGAnet")) install.packages("EGAnet")

# devtools::install_github("donaldRwilliams/BGGM", force = TRUE)
# devtools::install_github("donaldRwilliams/GGMnonreg", force = TRUE)
library(BGGM)
library(GGMnonreg)

require("tidyverse")
require("corrplot")
library(OpenMx)
library(devtools)
require(lavaan)

```
# Version 0.1 21.01.2020 - OAD
  
## Table of Contents
1.  Load libraries 
2.  Import and prepare data  & descriptives
3.  Compare different estimation techniques
3.1  PCL Network
4. Estiamte Networks incl. centrality and stability analyses
4.1 PCL Network
5. Confirmatory Network Analysis
5.1 PCL Network
6. Community Analysis
7. Predictability
8. Session info
9. Open Questions

<br/><br/>
<br/><br/>

## 2. Import and prepare data  & descriptives - this is the DSM-IV dataset (big data)
```{r}
# load data set
source('/home/or/Documents/va_data/readData.r')
```

#### Data cleanning and descriptive statistics
```{r}
# Addind time difference between PCL and PHQ
# gather info on both meds and no meda

# all patientes with PTSD and PCLTOT
pclAll <- dplyr::filter(vaDatclean, !is.na(BPCLTOT))
# plot pcl total score 
hist(pclAll$BPCLTOT)
# we have a minimum of 2 - so we have some NAs - let remove them
pclAll_Nas <- filter(pclAll, BPCLTOT <=16)
# total of 20 subjects with 16 or less in PCL (i.e. at least one missing variable)
# we can remove them from analysis
pclAll <- filter(pclAll, BPCLTOT >=17)
# 159577 patients
#pclNetwork <- pclNoNa # just medicated
pclNetwork <- pclAll
nrow(pclNetwork)
hist(pclNetwork$BPCLTOT)

pclNetwork$PCL_PHQdiff <- pclNetwork$PHQSURVEYDATE - pclNetwork$PCLSURVEYDATE
pclPHQNetwork <- filter(pclNetwork, PCL_PHQdiff <= 14 & PCL_PHQdiff >= 0) # removing patients with more than 14 apart between PHQ9 and PCL-M
pclPHQNetwork <- filter(pclPHQNetwork, BPCLTOT>=17)
hist(pclPHQNetwork$PCL_PHQdiff)
```

#### Sample descriptives of all subjects
```{r sample descriptives}
# gather info on both meds and no meda
# remove patients with more than 14 days apart PHQ and PCL
summary(pclAll$AGE_OCT01)
mean(pclAll$AGE_OCT01, na.rm=TRUE)
sd(pclAll$AGE_OCT01, na.rm=TRUE)
summary(pclAll$BPCLTOT)
mean(pclAll$BPCLTOT)
sd(pclAll$BPCLTOT)
table(pclAll$FEMALE)
summary(pclPHQNetwork$PHQSUMTOTAL)
mean(pclPHQNetwork$PHQSUMTOTAL, na.rm=TRUE)
sd(pclPHQNetwork$PHQSUMTOTAL, na.rm=TRUE)
table(pclPHQNetwork$FEMALE)

```

```{r}
# build data set only with PCL items
pclItems <- dplyr::select(pclAll, starts_with("PCL"))
pclPHQItems <- dplyr::select(pclPHQNetwork, starts_with("PCL"))

pclItems_noCluster <- dplyr::select(pclItems, -PCLFY, -PCLSURVEYDATE, -PCLRAWSCORE)
nrow(pclItems_noCluster)
pclItems_noCluster <- na.omit(pclItems_noCluster)
nrow(pclItems_noCluster)

```

## 3.  Compare different estimation techniques 
### 3.1  PCL Network
<br/><br/>
#### We compare several models to set the one's we would use thorugh the rest of analysis
#### First we do it on the PCL items alone (whole 150k)
```{r}
### A. Gaussian Graphical Model, regularized
df2 <- pclItems_noCluster

# Define labels
labels <- names(df2)

n2 <- estimateNetwork(df2, default="EBICglasso", corMethod = "cor", corArgs = list(method="spearman"), threshold=FALSE)
g2 <- plot(n2, legend.cex=.5, vsize=7)
# Severely skewed data, so we use Spearman over polychoric correlations here, as recommended (https://psycnet.apa.org/record/2018-13501-001)
# Warning: Dense network selected. But no negative edges, and bootstrapped edge weights look OK.
# However, we can still try threshold=TRUE as recommended; see next.

### B. Gaussian Graphical Model, regularized & thresholded
n3 <- estimateNetwork(df2, default="EBICglasso", corMethod = "cor", corArgs = list(method="spearman"), threshold=TRUE)
g3 <- plot(n3, layout=g2$layout, legend.cex=.5, vsize=7)

### C. Robustness: use new estimation procedure ggmModSelect (http://psychosystems.org/qgraph_1.5)
n4 <- estimateNetwork(df2, default="ggmModSelect", corMethod = "cor", corArgs = list(method="spearman"))
g4 <- plot(n4, layout=g2$layout, legend.cex=.5, vsize=7)

### D.1 Nonregularized network - I've used this one simply enough - please comment on better methods
n5 <- GGMregress(df2,IC = "BIC", method= "forward")
g5 <- qgraph(n5$pcor_or, layout = g2$layout, theme = "colorblind", labels = labels)

### D.2 - TOBIAS' SUGGESTION - Nonregularized network
n5_2 <- GGMnonreg::GGMboot(df2, alpha = 0.05, sims = 1000)
g5_2 <- qgraph(n5_2$mat_mean, layout = g2$layout, theme = "colorblind", labels = labels)

### E Bayesian Network
n6 <- explore(df2, prior_sd = 0.5, iter = 5000, cores = 4)
# H1 for this network is, that edges are not zero (either negative or positive), H0 is that they are zero #
E <- BGGM::select(n6, BF_cut = 10, alternative = "two.sided")
# Summary table of the network ##
summary(E, summarize = T, log = TRUE, digits = 2)  # log TRUE: BF= ln(BF), BF_10 = evidence for H1, BF_01 = evidence for H0
# Plot
g6 <- qgraph(E$partials_non_zero, legend = FALSE, layout = g2$layout, theme = "colorblind", labels = labels)

```

#### Correlate the different networks with the first network (EBIC, spearman) to check similarity
```{r}
cor(vechs(n2$graph), vechs(n4$graph)) # 0.997
cor(vechs(n2$graph), vechs(n4$graph), method="spearman") # 0.993
cor(vechs(n2$graph), vechs(n3$graph), method="spearman") #0.99
cor(vechs(n4$graph), vechs(n3$graph), method="spearman") #0.989
cor(vechs(n5$pcor_or), vechs(n2$graph), method = "spearman") #0.958
cor(vechs(n5_2$mat_mean), vechs(n2$graph), method = "spearman") #
cor(vechs(E$partials_non_zero), vechs(n2$graph), method = "spearman") #0.95

# Looks like all networks are highly correlated with each other. 
# We choose n2 from here on, alt
```

#### Estiamte Predictability (using mgm)
```{r}
fit1 <- mgm(data = na.omit(df2), type = rep('g', 17), lambdaSel = 'CV', level = rep(1,17), k = 3)
pred1 <- predict(fit1, na.omit(df2), errorCon = "R2")

pred1$errors #list with predcitability for each node

# Average node predictability #
R2_1<-as.numeric(pred1$errors$R2)
mean(R2_1) # 0.464

# ## Plot network and save as PDF 
g2 <- plot(n2, pie=R2_1, title="PCL-Network", legend.cex=.5, vsize=7)

```

<br/><br/>
<br/><br/>

## 4. Estiamte Networks incl. centrality and stability analyses
### 4.1 PCL Network
```{r network_pcl}
# building two kinds of group clusters. One (gr1) taken from Harpaz-Rotem, I., Tsai, J., Pietrzak, R. H., & Hoff, R. (2014).
gr1 <- list("Re-experiencing"=c(1:5), "Avoidance"=c(6:7), "Emotional numbing"=c(8:12),"Dysphoric arousal"=c(13:15), "Anxious arousal"=c(16:17)) #PTSD Clusters
gr_likeDSM <- list("Intrusion"=c(1:5), "Avoidance"=c(6:12), "Arousal"=c(13:17)) #PTSD symptoms categories B C D

g2 <- plot(n2, pie=R2_1, legend.cex=.5, vsize=7, theme = "colorblind", groups = gr1)
```

#### Centrality graph of PCL network:
```{r}
# sort by level of centrality (strength)
sort(centrality(n2)$OutDegree  , decreasing = T)
# plot centrality
centralityPlot(n2, include = "all")
```

#### Stability of PCL network
```{r bootstrap}
# Bootstrap 1:
boot1 <- bootnet(n2, nCores = 6, nBoots = 1000, type = "nonparametric") 
plot(boot1, labels = F, order = "sample")  + theme_minimal()
# now lets look at subjects
boot2 <- bootnet(n2, nCores = 6, nBoots = 1000, type = "case")
plot(boot2, labels = F, order = "sample") + theme_minimal()

#plot sig diff edges
plot(boot1, "edge", plot = "difference", onlyNonZero = TRUE, order = "sample")

#plot sig diff nodes
plot(boot1, "Strength") 
```


## 5. Confirmatory network analysis
#### We randomly sample 50% of population and run the model, then fit on the other 50%
<br/><br/>
<br/><br/>
### 5.1 PCL Network
#### Split Data into Training and Testing in R 
```{r}
sample_size = floor(0.5*nrow(pclItems_noCluster))
set.seed(777)

# randomly split data in r
picked = sample(seq_len(nrow(pclItems_noCluster)),size = sample_size)
train =pclItems_noCluster[picked,]
test =pclItems_noCluster[-picked,]

# Start run confirmatory analysis
# run model on half the subjects 
# shuold we consider comparing fit of different sets of models? or is it too much?
#net <-  estimateNetwork(train, default="EBICglasso", corMethod = "cor", corArgs = list(method="spearman"), threshold=FALSE)
net <- estimateNetwork(train, default = "ggmModSelect", verbose = FALSE)
network <- 1*(net$graph != 0)
model_frombootnet <- ggm(train, omega = network) %>% runmodel
```

#### Run analysis
```{r}
adjacency <- network #1*(getmatrix(model_frombootnet, "omega")!=0)
confirmatory <- ggm(test, omega = adjacency)
confirmatory <- confirmatory %>% runmodel

confirmatory %>% fit
```
```{r}
# saving to word table
library(rtf)
rtffile <- RTF("fitResults.doc")  # this can be an .rtf or a .doc
addParagraph(rtffile, "This is the output of fit results:\n")
addTable(rtffile, as.data.frame((confirmatory %>% fit)))
done(rtffile)
```

#### Compare fit indices
```{r}
compare(train = model_frombootnet , test = confirmatory)
```

## 6. Community Analysis
#### Data driven clustering using EGA
```{r}
# Cluster only PCL
egaPCL_glasso<-EGA(df2, plot.EGA = TRUE, steps = 4,  model = "glasso")
egaPCL_TMFG<-EGA(df2, plot.EGA = TRUE, steps = 4,  model = "TMFG")
# plot ega gLASSO
plot(egaPCL_glasso, theme = "colorblind", layout =g2$layout)

pclCFA_glasso <- CFA(ega.obj = egaPCL_glasso,data = df2,plot.CFA = TRUE, estimator = "WLSMV")
# plot CFA gLASSO
plot(pclCFA_glasso, theme="colorblind")

pclCFA_TMFG <- CFA(ega.obj = egaPCL_TMFG,data = df2,plot.CFA = TRUE, estimator = "WLSMV")

bootEGA_pcl <- bootEGA(df2, 500, type = "resampling", ncores = 5)
plot(bootEGA_pcl)
# bootstrap results in the same structure
# run cfa on 5 factor model
```

### Only PCL has three clusters (when EGA load is set to 0.7 )
## We should compare the confirmatory analysis of PTSD (PCL) based on the EGA with one based on the 5-factor model

```{r}
# gr1 <- list("Re-experiencing"=c(1:5), "Avoidance"=c(6:7), "Emotional numbing"=c(8:12),"Dysphoric arousal"=c(13:15), "Anxious arousal"=c(16:17))
model_5Factor <- ' ReExperiencing =~ PCL1 + PCL2 + PCL3 + PCL4 + PCL5
Avoidance =~ PCL6 + PCL7
Numbing =~ PCL8 + PCL9 + PCL10 + PCL11 + PCL12
DysphoricArousal =~ PCL13 + PCL14 + PCL15
AnxiousArousal =~ PCL16 + PCL17 '
theoModel <- cfa(model_5Factor, data = df2, estimator = "WLSMV")

# fit measures of 5 factor model
fitMeasures(theoModel, c("chisq","df","pvalue","srmr","cfi","rmsea"))
# fit measures of EGA
fitMeasures(pclCFA_glasso$fit, c("chisq","df","pvalue","srmr","cfi","rmsea"))
# seems like 5 models has a bit better fit
fitMeasures(pclCFA_TMFG$fit, c("chisq","df","pvalue","srmr","cfi","rmsea"))

# seems like TMGF is the worst option

# Lets run Chisq to compare each model

lavTestLRT(theoModel,pclCFA_TMFG$fit)
lavTestLRT(theoModel,pclCFA_glasso$fit)
```
## Lets plot the theoretical modeling, just for the sake of plotting
```{r}
semPlot::semPaths(theoModel, what = "est", layout = "spring", theme = "colorblind")#, title = FALSE, curvePivot = TRUE)
```
## We can next compare it with the DSM-IV model (3 factors)
```{r}
#gr_likeDSM <- list("Intrusion"=c(1:5), "Avoidance"=c(6:12), "Arousal"=c(13:17)) #PTSD symptoms categories B C D
dsm_Model <- 'Intrusion =~ PCL1 + PCL2 + PCL3 + PCL4 + PCL5 
Avoidance =~ PCL6 + PCL7 + PCL8 + PCL9 + PCL10 + PCL11 + PCL12
Arousal =~ PCL13 + PCL14 + PCL15 + PCL16 + PCL17'

DSMModel <- cfa(dsm_Model, data = df2, estimator = "WLSMV")
fitMeasures(DSMModel, c("chisq","df","pvalue","srmr","cfi","rmsea"))
lavTestLRT(theoModel,DSMModel)
lavTestLRT(DSMModel,pclCFA_glasso$fit, method ="satorra.2000")
```
5-factor model outperform theoretical model. 


## 7. Session info
```{r}
session_info()
```

<br/><br/>
<br/><br/>
# 8. Open Questions
